---
title: "STAT/MATH 495: Problem Set 06"
author: "Caleb Brooks"
date: "2017-10-17"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
```





# Collaboration

Please indicate who you collaborated with on this assignment: 





# Setup

Define truth, which again we know for the purposes of this assignment, but in
practice we won't:

* the true function f(x) i.e. the signal
* the true epsilon i.e. the noise, which in this case is Normal$(0, sd=\sigma)$.
Hence the standard deviation $\sigma$ determines the amount of noise.

```{r}
f <- function(x) {
  x^2
}
sigma <- 0.3
```

This is the target point we'll be trying to predict: $(0.95, f(0.95)) = (0.95, 0.95^2) = (0.95, 0.9025)$, Thus, the test set is just `x=0.95`

```{r}
x0 <- 0.95
test_set <- data_frame(x=x0)
```

This function generates a random sample of size $n$; think of this as a "get new
data" function. Random in terms of both:

* (New) the predictor x (uniform on [0,1])
* the amount of noise $\epsilon$

```{r}
generate_sample <- function(f, n, sigma) {
  sample <- data_frame(
    x = runif(n = n, min = 0, max = 1),
    f_x = f(x),
    epsilon = rnorm(n = n, mean = 0, sd = sigma),
    y = f_x + epsilon
  )
  # Recall: We don't observe f(x) and epsilon, just (x, y)
  sample <- sample %>% 
    select(x, y, epsilon)
  
  return(sample)
}
```

Define

* The number $n$ of observations $(x_i, y_i)$ in each sample. In the handout,
$n=100$ to keep plots uncrowded. Here we boost to $n=500$
* Number of samples of size $n$ to consider

```{r}
n <- 500
n_sample <- 10000
```


# Computation

```{r}
set.seed(619)
df2 <- numeric(10000)
df99 <- numeric(10000)

for (i in 1:n_sample){
  new_data <- generate_sample(f, n, sigma)
  error <- new_data$y-(new_data$x)^2
  
  s1 <- smooth.spline(new_data$x, new_data$y, df = 2)
  s2 <- smooth.spline(new_data$x, new_data$y, df = 99)  

  df2[i] = predict(s1, .95)$y
  df99[i] = predict(s2, .95)$y
}

df2_sqe <- (df2-(.95^2))^2
df99_sqe <- (df99-(.95^2))^2

mse1 <- mean(df2_sqe) 
mse2 <- mean(df99_sqe)

var1 <- df2 %>%
  `-`(mean(df2)) %>% 
  `^`(2) %>% 
  sum() %>% 
  `/`(length(df2)-1)
  
var2 <- df99 %>%
  `-`(mean(df99)) %>% 
  `^`(2) %>% 
  sum() %>% 
  `/`(length(df99)-1)

bias_sq1 <- (mean(df2)-.95^2)^2
bias_sq2 <- (mean(df99)-.95^2)^2
  
sum1 <- bias_sq1+var1
sum2 <- bias_sq2+var2
```


# Tables

As done in Lec 2.7, for both

* An `lm` regression AKA a `smooth.splines(x, y, df=2)` model fit 
* A `smooth.splines(x, y, df=99)` model fit 

output tables comparing:

|    DF|         MSE|   bias_squared|        var| irreducible|   sum|
|-----:|-----------:|--------------:|----------:|-----------:|------------------------:|
|     2| `r I(mse1)`|`r I(bias_sq1)`|`r I(var1)`|`r I(sigma)`|`r I(sum1)`|
|    99| `r I(mse2)`|`r I(bias_sq2)`|`r I(var2)`|`r I(sigma)`|`r I(sum2)`|

where `sum = bias_squared + var + irreducible`. You can created cleanly formatted tables like the one above by piping a data frame into `knitr::kable(digits=4)`.




# Analysis

**Questions**:

1. Based on the topics covered in Lec 2.7, name one possible "sanity check" for your results. Name another if you can.
1. In **two** sentences or less, give a rough sketch of what the procedure would
be to get the breakdown of $$\mbox{MSE}\left[\widehat{f}(x)\right]$$ for *all*
$x$ in this example, and not just for $$\mbox{MSE}\left[\widehat{f}(x_0)\right]
= \mbox{MSE}\left[\widehat{f}(0.95)\right]$$.
1. Which of the two models would you choose for predicting the point of interest and why?



**Answers**:

1.The change in bias and variance as DF increases shows what we would intuitively expect - that bias decreases and variance increases.

1. a similar procedure, but with random selection of x values from the relevant domain

1. As it has substantially less error overall, I would choose the model with df = 99 for the purpose of predicting a single point. 
