---
title: "STAT/MATH 495: Problem Set 06"
author: "WRITE YOUR NAME HERE"
date: "2017-10-17"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
```





# Collaboration

Please indicate who you collaborated with on this assignment: 





# Setup

Define truth, which again we know for the purposes of this assignment, but in
practice we won't:

* the true function f(x) i.e. the signal
* the true epsilon i.e. the noise, which in this case is Normal$(0, sd=\sigma)$.
Hence the standard deviation $\sigma$ determines the amount of noise.

```{r}
f <- function(x) {
  x^2
}
sigma <- 0.3
```

This is the target point we'll be trying to predict: $(0.95, f(0.95)) = (0.95, 0.95^2) = (0.95, 0.9025)$, Thus, the test set is just `x=0.95`

```{r}
x0 <- 0.95
test_set <- data_frame(x=x0)
```

This function generates a random sample of size $n$; think of this as a "get new
data" function. Random in terms of both:

* (New) the predictor x (uniform on [0,1])
* the amount of noise $\epsilon$

```{r}
generate_sample <- function(f, n, sigma) {
  sample <- data_frame(
    x = runif(n = n, min = 0, max = 1),
    f_x = f(x),
    epsilon = rnorm(n = n, mean = 0, sd = sigma),
    y = f_x + epsilon
  )
  # Recall: We don't observe f(x) and epsilon, just (x, y)
  sample <- sample %>% 
    select(x, y, epsilon)
  
  return(sample)
}
```

Define

* The number $n$ of observations $(x_i, y_i)$ in each sample. In the handout,
$n=100$ to keep plots uncrowded. Here we boost to $n=500$
* Number of samples of size $n$ to consider

```{r}
n <- 500
n_sample <- 10000
```


# Computation

```{r}
new_data <- generate_sample(f, n_sample, sigma)
error <- new_data$y-(new_data$x)^2

spline1 <- smooth.spline(x = new_data$x, y = new_data$y, df = 2)
spline2 <- smooth.spline(x = new_data$x, y = new_data$y, df = 99)

predict1 <- predict(spline1, new_data$x)
predict2 <- predict(spline2, new_data$x)

new_data$predict1 <- predict1$y
new_data$predict2 <- predict2$y

new_data$predict1error <- new_data$predict1 %>% 
  `-`((new_data$x)^2)
new_data$predict2error <- new_data$predict2 %>% 
  `-`((new_data$x)^2)

mse1 <- new_data$predict1error^2 %>% 
  sum %>%
  `/`(length(error)) %>% 
  round(5)
mse2 <- new_data$predict2error^2 %>% 
  sum %>%
  `/`(length(error)) %>% 
  round(5)

var1 <- new_data$predict1 %>% 
  `-`(mean(new_data$predict1)) %>% 
  `^`(2) %>% 
  sum %>%
  `/`(length(new_data$predict1)-1) %>% 
  round(5)
var2 <- new_data$predict2 %>% 
  `-`(mean(new_data$predict2)) %>% 
  `^`(2) %>% 
  sum %>%
  `/`(length(new_data$predict2)-1) %>% 
  round(5)

bias_sq1 <- (mse1*length(error))-var1-sigma
bias_sq2 <- (mse2*length(error))-var2-sigma
```


# Tables

As done in Lec 2.7, for both

* An `lm` regression AKA a `smooth.splines(x, y, df=2)` model fit 
* A `smooth.splines(x, y, df=99)` model fit 

output tables comparing:

|    DF|         MSE|   bias_squared|        var| irreducible|   sum|
|-----:|-----------:|--------------:|----------:|-----------:|------------------------:|
|     2| `r I(mse1)`|`r I(bias_sq1)`|`r I(var1)`|`r I(sigma)`|`r I(mse1*length(error))`|
|    99| `r I(mse2)`|`r I(bias_sq2)`|`r I(var2)`|`r I(sigma)`|`r I(mse2*length(error))`|

where `sum = bias_squared + var + irreducible`. You can created cleanly formatted tables like the one above by piping a data frame into `knitr::kable(digits=4)`.




# Analysis

**Questions**:

1. Based on the topics covered in Lec 2.7, name one possible "sanity check" for your results. Name another if you can.
1. In **two** sentences or less, give a rough sketch of what the procedure would
be to get the breakdown of $$\mbox{MSE}\left[\widehat{f}(x)\right]$$ for *all*
$x$ in this example, and not just for $$\mbox{MSE}\left[\widehat{f}(x_0)\right]
= \mbox{MSE}\left[\widehat{f}(0.95)\right]$$.
1. Which of the two models would you choose for predicting the point of interest and why?



**Answers**:

1.The change in bias and variance as DF increases shows what we would intuitively expect - that bias decreases and variance increases.

1. I think this is actually what I did - will change for resubmission if need be.

1. As it has substantially less error overall, I would choose the model with df = 99
